---
title: "p8105_hw1_eej2124"
author: "Emma Johnston"
date: "2026-09-19"
output: github_document
---

```{r message = F}
library(tidyverse)
```

## Problem 1

```{r}
data("penguins", package = "palmerpenguins")
```

This data set includes information on the `r colnames(penguins)`. It has `r ncol(penguins)` columns and `r nrow(penguins)` rows. The mean flipper length is `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), 2)` mm. 

# How do I include "values" of important variables?

# Had to remove NA from df - all good?

### Scatterplot 

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
```

# Should I clean up this scatterplot - axis titles, etc?


```{r}
ggsave("scatterplot.png")
```

## Problem 2 

```{r}
df = 
  tibble(
    vec_numeric = rnorm(10),
    vec_logical = vec_numeric > 0, 
    vec_character = character(length = 10),
    vec_factor = factor(c(length = 10), levels = 3)
  )
```
# Fix vector factor 
# Why are means so low?
# Should I set.seed?

The mean of the numeric vector is `r round(mean(pull(df, vec_numeric), 2))`. The mean of the logical vector is `r round(mean(pull(df, vec_logical), 2))`. The mean of the character vector is `r round(mean(pull(df, vec_character), 2))`. The mean of the factor vector is `r round(mean(pull(df, vec_factor), 2))`. You cannot take the mean of neither the character nor factor vectors. 

# Write as.numeric code 




